Followers Are Not Enough: Beyond Structural Communities in Online Social Networks \\Followers Are Not Enough: A Question-Oriented Approach to Community Detection

### 2014-07-31 paper-for_arxiv/methods.tex
Addressed revisions from next actions.

The dataset for this study consisted of the tweets of 15,000 Twitter users over a 9 week period (from April 25th to June 25th 2011). The users are embedded in a network collected by performing an intelligent breadth-first expansion from a random seed user. In particular, once the seed user was chosen, the network was expanded to include his/her followers, but only included users considered to be `active' (i.e., users who tweeted at least once per day over the past one hundred days). Network collection continued in this fashion by considering the active followers of the active followers of the seed, and so on until 15,000 users were added to the network.

@@

 We computed the transfer entropy on each coarsened time series with lag $k$ ranging from 1 to 6, this corresponds to a lag of ten minutes to an hour. The choice of lag must balance a trade-off between additional information and sparsity of samples: as we increase the lag $k$, we account for longer range dependencies, but we also decrease the number of samples available to infer a higher dimensional predictive distribution, see Appendix A for a discussion. Ultimately, due to similarities in the underlying communities we chose a lag-4 transfer entropy. All references to activity-based weights, unless otherwise noted, refer to this choice of lag.

>>>>

The dataset for this study consisted of the tweets of 15,000 Twitter users over a 9 week period (from April 25th to June 25th 2011). The users are embedded in a network collected by performing an intelligent breadth-first expansion from a random seed user. In particular, once the seed user was chosen, the network was expanded to include his/her followers, but only included users considered to be `active' (i.e., users who tweeted at least once per day over the past one hundred days). Network collection continued in this fashion by considering the active followers of the active followers of the seed, and so on until 15,000 users were added to the network. The only biasing steps in this procedure are the selection of the seed node and the filtering out of inactive users. However, because all of the community types we are interested in involve some sort of interaction between users, each other, and their content, the procedure provides a relatively question-agnostic follower-followee network for use with the different question-based analyses.

@@

 We computed the transfer entropy on each coarsened time series with lag $k$ ranging from 1 to 6, this corresponds to a lag of ten minutes to an hour. The choice of lag must balance a trade-off between additional information and sparsity of samples: as we increase the lag $k$, we account for longer range dependencies, but we also decrease the number of samples available to infer a higher dimensional predictive distribution, see Appendix A for a discussion. As we will see, the underlying communities resulting from the different lags have similar structure.

### 2014-07-31 paper-for_arxiv/results.tex
Addressed revisions from next actions.

We see in Table~\ref{Table-comm_count} that the topic- and interaction-based communities have the most singletons, with the activity-based community dominating this measure. This result for the activity-based community is partially an artifact of a property of the retweet/mention weighting: 717 of the users were disconnected from the network by how the weights were defined, resulting in `orphan' nodes which we have included in the collection of singletons for all of our analyses. However, even after accounting for this artifact, the interaction-based network still has the most non-orphan singletons. This seems to indicate that a large fraction of the 6917 (nearly 25\%) do not interact with each other in a concerted way that would mark them as a community under our interaction-based definition. This agrees with a result previously reported in~\cite{romero2011influence} about how most users passively interact with incoming information on Twitter.

@@

Next we consider the distribution of community sizes across the community types. The complementary cumulative distribution of community sizes is given in Figure~\ref{Fig-community_size_distribution}. Note that both axes are plotted on log-scales. Thus, for a fixed community size $s$, Figure~\ref{Fig-community_size_distribution} shows the proportion of communities of size greater than $s$ for each community type. We see that the community distributions have longer tails for the non-structural networks, and that the interaction-based network has the longest tail. The largest communities for the structural, activity-based, topic-based, and interaction-based networks have 198, 358, 338, and 811 members, respectively. Most importantly, we see that the distributions of community sizes differ across the community types, highlighting that the different networks give rise to different large-scale community structure dependent on the particular weighting of the structural network.

@@

We considered the normalized mutual information between the communities inferred from the structural network and the networks weighted with lag 1 through 6 transfer entropies, hashtag similarity, and mention, retweet, and mention-retweet activity. The resulting $\text{NMI}(\mathcal{C}_{i}, \mathcal{C}_{j})$ are shown in Figure~\ref{Fig-compare_coverings}. We see that similarity between the coverings is dictated by the generic community type (structural, activity-based, etc.). That is, the transfer entropy coverings are more similar to each other than to any of the other coverings, with a similar result for the mention, retweet, and mention-retweet coverings. Interestingly, the coverings resulting from the different weightings are all more similar to each other than to the structural covering from the unweighted network. Also note that the covering based on the hashtag similarities are different from all of the other weight-based coverings.

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{edge-types.pdf}
	\caption{A schematic of how, given a covering, the edges of the network can be partitioned using \ref{Eqn-edge_types} into inter-edges, intra-edges, and mixed-edges. Inter-edges (red) cross community boundaries. Intra-edges (blue) remain inside community boundaries. Mixed-edges (purple) both remain in and cross community boundaries due to overlap in community membership. Each column corresponds to the same collection of weights, but partitioned using a different covering. Each row corresponds to the same covering, but for the different weights.}
	\label{Fig-edge_types}
\end{figure}

@@

Each community type (Structural, Activity-based, Topic-based, Interaction-based) induces a different partition of the edges, and each edge type (Transfer Entropy, Hashtag, Mention-Retweet) induces a different distribution of weights. That is, let $W_{u \to f}$ be the weight on a randomly selected edge $E_{u \to f}$ from the network, and compute
\begin{align}
	P(W_{u \to f} \leq w | T(E_{u \to f}) = t),
\end{align}
the empirical distribution over the edge weights conditioned on $t$ being one of inter, intra, and mixed. For meaningful community structure, we expect the distribution of edge weights to differ with the edge type $t$. If the community structure were arbitrary, the edge weights would be independent of the edge type, and all three conditional distributions would be identical.
The densities associated with these distributions are shown in Figure~\ref{Fig-distributions_by_types}.
\begin{figure}[!h]
	\centering
	\includegraphics[width=1\textwidth]{densities-edges.eps}
	\caption{The density of edge weights for different community types (rows) and weight types (columns). The red,  blue, and magenta values below each collection of densities indicate the median of weight on intra-, inter-, and mixed-edges, respectively.}
	\label{Fig-distributions_by_types}
\end{figure}

Each \emph{column} of Figure~\ref{Fig-distributions_by_types} demonstrates how the density of weights change with the community type used to induce the partition. For example, the second column shows how the density of hashtag weights change based on the community partition. In general, the densities differ in non-trivial ways across the edge types and coverings. However, we see that the medians of the distributions provide a useful summary statistic. Under each collection of density functions, we list the median value of the inter, intra, and mixed-edges in red, blue, and magenta, respectively. Unsurprisingly, we see that the greatest difference between the distributions occurs when we consider the topic-based communities, since the topic-based communities are determined based on the hashtag weights. However, we also see differences in the distributions when the Community-type / Weight-type pairs do not match. For example, for the covering corresponding to the activity-based communities, the hashtag weights tend to be higher for edges internal to communities than between communities. Similarly, for the covering corresponding to the topic-based communities, the transfer entropy weights tend to be higher on edges within (intra) and between (inter) communities, and lower for edges between members with multiple, non-identical memberships (mixed).

>>>>

We see in Table~\ref{Table-comm_count} that the topic- and interaction-based communities have the most singletons. This result for the interaction-based community is partially an artifact of the retweet/mention weighting: 717 of the users were disconnected from the network by how the weights were defined, resulting in `orphan' nodes which we have included in the collection of singletons for all of our analyses. However, even after accounting for this artifact, the interaction-based network still has the most non-orphan singletons. This seems to indicate that a large fraction of the 6917 (nearly 25\%) do not interact with each other in a concerted way that would mark them as a community under our interaction-based definition. This agrees with a result previously reported in~\cite{romero2011influence} about how most users passively interact with incoming information on Twitter.

@@

Next we consider the distribution of community sizes across the community types. The complementary cumulative distribution of community sizes is given in Figure~\ref{Fig-community_size_distribution}. Note that both axes are plotted on log-scales. Thus, for a fixed community size $s$, Figure~\ref{Fig-community_size_distribution} shows the proportion of communities of size greater than $s$ for each community type. We see that the community distributions have longer tails for the non-structural networks, and that the interaction-based network has the longest tail.
%The largest communities for the structural, activity-based, topic-based, and interaction-based networks have 198, 358, 338, and 811 members, respectively.
Note that the activity-based communities, using transfer entropy estimated with varying lags, seem to converge on a similar large-scale community structure around lag 3. That is, the communities based on lag 1 and lag 2 transfer entropies are generally larger, and these communities resolve into smaller communities as the lag increases beyond 2. The distribution of community sizes for lag larger than 2 is insensitive to the lag.
Most importantly, we see that the distributions of community sizes differ across the community types, highlighting that the different networks give rise to different large-scale community structure dependent on the particular weighting of the structural network.

@@

We considered the normalized mutual information between the communities inferred from the structural network and the networks weighted with lag 1 through 6 transfer entropies, hashtag similarity, and mention, retweet, and mention-retweet activity. The resulting $\text{NMI}(\mathcal{C}_{i}, \mathcal{C}_{j})$ are shown in Figure~\ref{Fig-compare_coverings}. We see that similarity between the coverings is dictated by the generic community type (structural, activity-based, etc.). That is, the transfer entropy coverings are more similar to each other than to any of the other coverings, with a similar result for the mention, retweet, and mention-retweet coverings. Interestingly, the coverings resulting from the different weightings are all more similar to each other than to the structural covering from the unweighted network. Also note that the covering based on the hashtag similarities are different from all of the other weight-based coverings. In agreement with the results reported in the previous section, we see that the activity-based communities share similar structure for lags greater than 2. Because of these two results, in the remainder of the paper, we focus on the activity-based communities inferred using the lag-4 transfer entropy.

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{edge-types.pdf}
	\caption{A schematic of how, given a covering, the edges of the network can be partitioned using (\ref{Eqn-edge_types}) into inter-edges, intra-edges, and mixed-edges. Inter-edges (red dashed) cross community boundaries. Intra-edges (blue solid) remain inside community boundaries. Mixed-edges (purple dotted) both remain in and cross community boundaries due to overlap in community membership. Each column corresponds to the same collection of weights, but partitioned using a different covering. Each row corresponds to the same covering, but for the different weights.}
	\label{Fig-edge_types}
\end{figure}

@@

Each community type (Structural, Activity-based, Topic-based, Interaction-based) induces a different partition of the edges, and each edge type (Transfer Entropy, Hashtag, Mention-Retweet) induces a different distribution of weights. That is, let $W_{u \to f}$ be the weight on an edge $E_{u \to f}$ chosen uniformly at random from the network, and compute
\begin{align}
	P(W_{u \to f} \leq w_{u \to f} | T(E_{u \to f}) = t),
\end{align}
the empirical distribution over the edge weights conditioned on the edge type $t$ being one of inter, intra, and mixed. For meaningful community structure, we expect the distribution of edge weights to differ with the edge type $t$. If the community structure were arbitrary, the edge weights would be independent of the edge type, and all three conditional distributions would be identical.
The densities associated with these distributions are shown in Figure~\ref{Fig-distributions_by_types}.
\begin{figure}[!h]
	\centering
	\includegraphics[width=1\textwidth]{densities-edges.eps}
	\caption{The density of edge weights for different community types (rows) and weight types (columns). The red,  blue, and purple values below each collection of densities indicate the median of weight on intra-, inter-, and mixed-edges, respectively.}
	\label{Fig-distributions_by_types}
\end{figure}

Each \emph{column} of Figure~\ref{Fig-distributions_by_types} demonstrates how the density of weights change with the community type used to induce the partition. For example, the second column shows how the density of hashtag weights change based on the community partition. In general, the densities differ in non-trivial ways across the edge types and coverings. However, we see that the medians of the distributions provide a useful summary statistic. Under each collection of density functions, we list the median value of the inter, intra, and mixed-edges in red, blue, and purple, respectively. Unsurprisingly, we see that the greatest difference between the distributions occurs when we consider the topic-based communities, since the topic-based communities are determined based on the hashtag weights. However, we also see differences in the distributions when the Community-type / Weight-type pairs do not match. For example, for the covering corresponding to the activity-based communities, the hashtag weights tend to be higher for edges internal to communities than between communities. Similarly, for the covering corresponding to the topic-based communities, the transfer entropy weights tend to be higher on edges within (intra) and between (inter) communities, and lower for edges between members with multiple, non-identical memberships (mixed).
